---
title: "Visualisation_kongeraekken"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2023-10-03"
---
# Download packages and read the csv file containing kongerækken
```{r}
library(tidyverse)
kongeraekken <- read_csv("data_homework38/Kongerækken - Ark1-kopi.csv")

```
# Finding duration of reign
To find duration of reign i used Mutate to subtract start of reign from end of reign
Then i used the function glimps() to see a glimpse of how the spreadsheet looked with the new column duration of rule
```{r}

kongeraekken %>% 
  mutate(duration_of_rule = end_of_reign - start_of_reign) %>% 
  glimpse()

```
# Visualisation
to visualise I used ggplot and the geom_line and geom_smooth.
I choose the theme minimal which made the background white.
I Used labs() to give the visualisation a title, and give names to the x and y axes.
When doing so i am given two warnings and they both relate to the fact that Gorm den Gamles start of reign is a missing value (NA) and therefore R hasn't been able to calculate Gorm den Gamles duration of rule and therefore can't visualise it either

```{r}

kongeraekken %>% 
  ggplot(aes(x = end_of_reign, y = duration_of_rule)) +
  geom_line() +
  geom_smooth() + 
  theme_minimal() +
  labs( title = "Duration of reign among Danish rulers",
        x = "end of reign",
        y = "duration of reign")
```


