---
title: "Are we more civilized today?"
author: "Clara Sydow"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE)
```
This exercise is based on the dataset provided by [OurWorldInData project](https://ourworldindata.org/homicides) based at the Oxford University.

## The long-term trend in Homicides in Western Europe

Understanding how homicide rates have changed prior to the modern era requires the help of historians and archivists. Manuel Eisner, a criminology professor at the University of Cambridge, and his colleagues published the Historical Violence Database \href{https://cjrc.osu.edu/research/interdisciplinary/hvd}: a compilation of data on long-term trends in homicide rates, in addition to qualitative information such as the cause of death, perpetrator and victim. This database is limited to countries with relatively complete historical records on violence and crime – mainly Western Europe and the US.

Starting in the second half of the nineteenth century, these European regions have consistent police records of those accused of murder or manslaughter and annual counts of homicide victims. To go back further in time, reaching as far back as the thirteenth century, Eisner collected estimates (from historical records of coroner reports, court trials, and the police) of homicide rates made in over ninety publications by scholars.

Homicide rates – measured as the number of homicides per 100,000 individuals – up to 1990 are sourced from Eisner’s (2003) publication and the Historical Violence Database. 

Are homicide rates in Europe today lower or higher than in the past? Using the provided dataset, display and describe the long-run homicide rates for the five European regions: Italy, England, Germany, Netherlands and Scandinavia.

```{r library}
library(tidyverse)
```

## Load the available data from ourworldindata.org

You should always interrogate the source of your data. Who compiled it, from where, what is missing, how representative the data are? Check the data/Metadata.txt to learn about the data provenance.

```{r load-data}
Western_Europe <- read_csv("data/homicide-rates-across-western-europe.csv")
```

## Inspect the data

How clean and analysis-ready is the dataset? Do you understand what the column names represent? What is the difference between rate and homicide number?


```{r inspect}
head(Western_Europe)
```

Ok, the data look good except for the column `Homicide rate in Europe over long-term (per 100,000)` which is not very easy to work with.

- Use the `names()` function and assignment key to relabel this column to `homicides_per_100k` 

```{r relabel column}
names(Western_Europe)
colnames(Western_Europe)[4] = "homicides_per_100k"
```

Now, that you have looked at what the data looks like and what it represents, and streamlined it, let's see what big picture it contains.

## Let's see what the long-term trend is in homicides

- use `ggplot()` function and remember the`+` at the end of the line
- chose a meaningful `geom_......()`  for geometry (hint: points are not great)
- load `Year` on the `x` axis and `homicides_per_100k` column in y axis
- to color individual country entries consistently, assign the country column to the argument `color`.
- provide meaningful title and axis labels
- remember to change the `eval` flag so that the code chunk renders when knitted 

```{r plot-homicides, eval = TRUE}
Western_Europe %>% 
  ggplot(aes(x = Year, y = homicides_per_100k)) + 
  geom_smooth(aes(fill = Entity)) +
  labs( title = "The long-term homicides trend",
        x = "Years",
        y = "Homicides per 100,000")

```
Alright, the homicide rates should all be descending over time. What a comfort. But the viz is not super clear. Let's check the rates for individual countries.

## Uncouple the homicides of individual countries for easier view

You can visualize each country's trend separately by adding an extra argument to the ggplot, the `facet_wrap()` and feeding it the country column. If in doubt, check your ggplot tutorial and your country column name for exact usage.

- reuse the ggplot from the chunk above
- insert `facet_wrap()` after the specification of geometry to split countries in separate charts
- change the facet "layout" to two columns and three rows so that the trends are easier to see in horizontal layout. 


```{r facet_wrap_function, eval = TRUE}

Western_Europe %>% 
  ggplot(aes(x = Year, y = homicides_per_100k)) + 
  geom_smooth(aes(fill = Entity)) +
  facet_wrap(~ Entity, nrow = 3, ncol = 2) +
  labs( title = "The long-term homicides trend in different countries",
        x = "Years",
        y = "Homicides per 100,000") +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(title = "Country"))
  
```

## Compare the trends in homicide with the pattern of reign duration among Danish rulers through time.

- Load your Danish king dataset. Hopefully it is tidy and your years and duration of reign are all numeric.
- You need to have a consistent way of plotting the rulers' reign on the x axis, so I recommend you create a midyear column by calculating the middle of each monarch's rule (Hint: `midyear` = `endyear` - (`endyear`-`startyear`)/2)
- Start a ggplot plotting midyear on x axis and duration on y axis
- Try `geom_smooth()` for geometry 
- Provide meaningful labels and a title
- How would you characterize the trend compared to the homicides above?

### How would you characterize the trend compared to the homicides above?

It seems like while the duration of reign among Danish rulers gets longer, the number of homicides per 100.000 gets lesser the closer to the year 2000 you get

#### Warning: 
The warning about remowing 1 row is because Gorm den Gamles start of reign was NA so it couldn't findt the duration of rule or midyear


```{r kongeraekken, eval = TRUE}

kongeraekken <- read.csv("data/Kongerækken - Ark1-kopi 2.csv")


kongerækken_plotting<- kongeraekken %>% 
  mutate(duration_of_rule = end_of_reign - start_of_reign) %>%
  mutate(midyear = end_of_reign - (end_of_reign - start_of_reign)/2)

kongerækken_plotting %>% 
  ggplot(aes(x = midyear, y = duration_of_rule)) +
  geom_smooth() + 
  labs( title = "Reign duration among Danish rulers",
        x = "The middle of the monarch's rule",
        y = "Duration of rule")

```


## Final Tasks

1) Plot: In the facetted plot above, move the legend from the current position on the side to below the facets, and label it "Country" instead of "Entity".

2) Rmarkdown: 

  - edit the author of the document, and convert 'Final Tasks' into heading #2 (like the other headings) 
  - add a `floating table of contents` to your Rmarkdown document, 
  - provide informative `chunk-names` and edit flags in your R chunks, and
  - automatically generate a `timestamp` to show when the document was last updated. (Hint: check the Rmarkdown episode in our Data Carpentry tutorial)

3) Question: In <250 words articulate your answer on the basis of the data visualisations to the following question: are we more civilized today?

### Answer to question nr 3: are we more civilized today?
On the basis of the data visualisations made in this text I want to articulate an answer to the question: Are we more civilized today? The three data visualisations in this homework assignment are 1) the long-term homicides trends (in Europe), 2) the long-term homicides trends in different countries (in Europe), 3) reign duration among Danish rulers. The two first visualisations show how the number of homicides per 100.000 are decreasing as we get closer to the present, which means that fewer people are getting murdered and fewer people are murdering another human being. This could be a sign of a less barbaric and more civilized society. The visualisation about the Danish rulers shows us that the duration of their reigns are getting longer the closer we get to the present, but I can't really think of a way to interpret it as a sign of us being more civilized today. Overall I don't think the visualisations in this homework assignment gets me even close to being able to articulate a fair answer to the question. I believe the data in the visualisations are shown as if they are written in stone and thereby are facts. And I think the question being asked can be interpreted in many different ways and can be answered by looking at many different aspects about society. The question also contains a lot of components that can be interpreted differently: who are "we"? What does "civilized'' mean? When are "we more civilized" than other times? So I believe these visualisations are only aspects of what could lead to a fair answer to the question. 
