---
title: "Game Of Thrones Sentiment Analysis"
author: Clara Katrine Sydow
date: 'Created on 6 october 2023 and updated `r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(here)

# For text mining:
library(pdftools)
library(tidytext)
library(textdata) 
library(ggwordcloud)
```

#Loading the data, an Pdf-file
```{r loading data}
got_path <- here("data", "got.pdf")
got_text <- pdf_text(got_path)
```


# Single page
```{r single-page}
got_p502 <- got_text[502]
got_p502
```
## Comparison with the Pdf-file
In comparison with page 502 in the Pdf-file, R has added \n whenever there's a new line

# Wrangling
```{r split-lines}
got_df <- data.frame(got_text) %>% 
  mutate(text_full = str_split(got_text, pattern = '\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) 

```

# Get the tokens
```{r tokenize}
got_tokens <- got_df %>% 
  unnest_tokens(word, text_full)
got_tokens

```
# Count the words

```{r count-words}
got_wc <- got_tokens %>% 
  count(word) %>% 
  arrange(-n)

got_wc
```

# Remove stop words
We will *remove* stop words using `tidyr::anti_join()`:
```{r stopwords}
got_stop <- got_tokens %>% 
  anti_join(stop_words) %>% 
  select(-got_text)
```
Now check the counts again: 
```{r count-words2}
got_swc <- got_stop %>% 
  count(word) %>% 
  arrange(-n)
```

How to get rid of all the numbers (non-text) in `got_stop`?
```{r skip-numbers}
got_no_numeric <- got_stop %>% 
  filter(is.na(as.numeric(word)))

got_no_numeric
```

# A word cloud of Game of Thrones words (non-numeric)

```{r wordcloud-prep}
# There are 11209 unique words 
length(unique(got_no_numeric$word))

# We don't want to include all the words cloud. Let's filter to only include the top 100 most frequent?
got_top100 <- got_no_numeric %>% 
  count(word) %>% 
  arrange(-n) %>% 
  head(100)

#To show the top 100 words:
got_top100
```
```{r wordcloud}
got_cloud <- ggplot(data = got_top100, aes(label = word)) +
  geom_text_wordcloud() +
  theme_minimal()

got_cloud
```
Let's customize it a bit:
```{r wordcloud-pro}
ggplot(data = got_top100, aes(label = word, size = n)) +
  geom_text_wordcloud_area(aes(color = n), shape = "diamond") +
  scale_size_area(max_size = 12) +
  scale_color_gradientn(colors = c("darkgreen","blue","red")) +
  theme_minimal()
```
# Sentiment analysis
The three general-purpose lexicons are

  -  AFINN from Finn Årup Nielsen,
  -  bing from Bing Liu and collaborators, and
  -  nrc from Saif Mohammad and Peter Turney
  
"afinn": Words ranked from -5 (very negative) to +5 (very positive)
```{r afinn, loading the lexicons}
get_sentiments(lexicon = "afinn")
# Note: may be prompted to download (yes)

# Let's look at the pretty positive words:
afinn_pos <- get_sentiments("afinn") %>% 
  filter(value %in% c(3,4,5))

afinn_pos
```

bing: binary, "positive" or "negative"
```{r bing}
get_sentiments(lexicon = "bing")
```

nrc:https://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm
Includes bins for 8 emotions (anger, anticipation, disgust, fear, joy, sadness, surprise, trust) and positive / negative. 

Now nrc:
```{r nrc}
get_sentiments(lexicon = "nrc")
```

Let's do sentiment analysis on the Game of Thrones book data using afinn, and nrc.
## Sentiment analysis with afinn: 

First, bind words in `got_stop` to `afinn` lexicon:
```{r bind-afinn}
got_afinn <- got_stop %>% 
  inner_join(get_sentiments("afinn"))
```

Let's find some counts (by sentiment ranking):
```{r count-afinn}
got_afinn_hist <- got_afinn %>% 
  count(value)

# Plot them: 
ggplot(data = got_afinn_hist, aes(x = value, y = n)) +
  geom_col()
```
## About the graph
There are mostly negative words in the book according to the graph

Investigate some of the words in a bit more depth:
```{r afinn-2}
# What are these '2' words?
got_afinn_m5 <- got_afinn %>% 
  filter(value == -5)
```

```{r afinn-2-more}
# Check the unique 2-score words:
unique(got_afinn_m5$word)

# Count & plot them
got_afinn_m5_n <- got_afinn_m5 %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = fct_reorder(factor(word), n))


ggplot(data = got_afinn_m5_n, aes(x = word, y = n)) +
  geom_col() +
  coord_flip()
```
Or we can summarize sentiment for the report: 
```{r summarize-afinn}
got_summary <- got_afinn %>% 
  summarize(
    mean_score = mean(value),
    median_score = median(value)
  )

```

## Summarizing

The mean score is -0.5419666
The median score is -1

The mean and median clearly indicate negative overall sentiments in the Game of Thrones book.

## NRC lexicon for sentiment analysis

We can use the NRC lexicon to start "binning" text by the feelings they're typically associated with. As above, we'll use inner_join() to combine the GOT non-stopword text with the nrc lexicon: 

```{r bind-bing}
got_nrc <- got_stop %>% 
  inner_join(get_sentiments("nrc"))
```

## We should check which words are excluded using `anti_join()`:

```{r check-exclusions}
got_exclude <- got_stop %>% 
  anti_join(get_sentiments("nrc"))

# View(ipcc_exclude)

# Count to find the most excluded:
got_exclude_n <- got_exclude %>% 
  count(word, sort = TRUE)

head(got_exclude_n)
```
## Now find some counts: 
```{r count-bing}
got_nrc_n <- got_nrc %>% 
  count(sentiment, sort = TRUE)

# And plot them:

ggplot(data = got_nrc_n, aes(x = sentiment, y = n)) +
  geom_col()
```

## Or count by sentiment *and* word, then facet:
```{r count-nrc}
got_nrc_n5 <- got_nrc %>% 
  count(word,sentiment, sort = TRUE) %>% 
  group_by(sentiment) %>% 
  top_n(5) %>% 
  ungroup()

got_nrc_gg <- ggplot(data = got_nrc_n5, aes(x = reorder(word,n), y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, ncol = 2, scales = "free") +
  coord_flip() +
  theme_minimal() +
  labs(x = "Word", y = "count")

# Show it
got_nrc_gg

# Save it
ggsave(plot = got_nrc_gg, 
       here("figures","got_nrc_sentiment.png"), 
       height = 8, 
       width = 5)

```

Wait, so "lord" is showing up in NRC lexicon as "disgust"? Let's check:
```{r nrc-confidence}
lord <- get_sentiments(lexicon = "nrc") %>% 
  filter(word == "lord")

# Yep, check it out:
lord
```

  




